---
title: "HW 1 Pt. 2"
author: "Ali Banijamali"
---

* **First**, load the file *M01_quasi_twitter.csv*


```{r}
twitter <- read.csv(file = "C:/Users/Ali/Desktop/DSCS 6030 - Intro to data Mining and Machine Learning/Module 1/Assignment/M01_quasi_twitter.csv",
         header=TRUE, sep=",")
```

* **Next**, answer the following questions for the data in each column:

    * How is the data distributed?
    
    * What happens when you test distribution assumptions (e.g. normal distributions or skewed, etc)?
    
    * What are the summary statistics?
    
    * Are there anomalies/outliers?
    
    * Can you identify the following:
    
        * useful raw data and transforms (e.g. log(x))
        
        * data quality problems
        
        * outliers
        
        * subsets of interest
        
    * Finally, suggest any functional relationships.

> Here is the summary of all the data:

```{r, echo=FALSE}

lapply(twitter, summary) 

```

> Here are the histograms of the numerical data:

```{r, echo=FALSE}

hist(twitter$created_at_month, breaks=10, col="1")

hist(twitter$created_at_day, breaks=10, col="2")

hist(twitter$created_at_year, breaks=10, col="3")

hist(twitter$friends_count, breaks=10, col="4")

hist(twitter$followers_count, breaks=10, col="5")

hist(twitter$statuses_count, breaks=10, col="6")

hist(twitter$favourites_count, breaks=10, col="7")

hist(twitter$favourited_count, breaks=10, col="8")

hist(twitter$dob_day, breaks=10, col="9")

hist(twitter$dob_year, breaks=10, col="10")

hist(twitter$dob_month, breaks=10, col="11")

hist(twitter$mobile_favourites_count, breaks=10, col="12")

hist(twitter$mobile_favourited_count, breaks=10, col="13")

hist(twitter$education, breaks=10, col="14")

hist(twitter$experience, breaks=10, col="15")

hist(twitter$age, breaks=10, col="16")

hist(twitter$wage, breaks=10, col="17")

hist(twitter$retweeted_count, breaks=10, col="18")

hist(twitter$retweet_count, breaks=10, col="19")

hist(twitter$height, breaks=10, col="20")

```



* created at month/day/year are almost uniform.

* education, experience, age, wage and height are almost normal.

* For distinction between skewed and normal we have to look at the median and mean in the data summaries:

* retweet count, mobile favourited count, mobile favourites count, favourited count, favourites count, statuses count, followers count, friends count are all skewed distributions.

* By looking at the countries: Kuwait and Portugal have the least number of twits (49) and USA has the most number of twits (14905).

* By looking at the locations: SoCal has the least frequency (37) and Mexico has the most frequency (122)

* By looking at the gender: distribution we can see that there are more men (14569) on twitter compared to women (7319).

* By looking at the races: Most of the users are white (18032) and the fewest number of users are indian (162).
 
















